\lecture{2}{07 Jan.\ 11:00}{Measure}
\begin{eg}
	We first see some examples.
	\begin{enumerate}[(a)]
		\item Let \(\mathcal{A} = \mathcal{P} (X)\), which is the power \hyperref[def:sigma-algebra]{\(\sigma\)-algebra}.
		\item Let \(\mathcal{A} = \{\varnothing , X\}\), which is a trivial \hyperref[def:sigma-algebra]{\(\sigma\)-algebra}.
		\item Let \(B\subset X\), \(B\neq \varnothing \), \(B\neq X\). Then we see that \(\mathcal{A} = \{\varnothing , B, B^{c}, X\}\) is a \hyperref[def:sigma-algebra]{\(\sigma\)-algebra}.
	\end{enumerate}
\end{eg}

\begin{lemma}
	Let \(\mathcal{A}_{\alpha}\), \(\alpha\in I\), be a family of \hyperref[def:sigma-algebra]{\(\sigma\)-algebra} on \(X\). Then \(\bigcap_{\alpha\in I} \mathcal{A}_{\alpha}\) is a \hyperref[def:sigma-algebra]{\(\sigma\)-algebra} on \(X\).
\end{lemma}
\begin{proof}
	A simple proof can be made as follows. Firstly, \(\varnothing \in \mathcal{A}_{\alpha}\) for every \(\alpha\) clearly. Moreover, closure under complement and countable unions for every \(\mathcal{A}_{\alpha}\) implies the same must be true for \(\bigcap_{\alpha\in I} \mathcal{A}_{\alpha}\). Hence, \(\bigcap_{\alpha\in I} \mathcal{A}_{\alpha}\) is a \hyperref[def:sigma-algebra]{\(\sigma\)-algebra}.
	\begin{remark}
		Notice that \(I\) may be an uncountable intersection.
	\end{remark}
\end{proof}

The above allows us to give the following definition.
\begin{definition}[Generation of \(\sigma\)-algebra]\label{def:generation-of-sigma-algebra}
	Given \(\mathcal{E} \subset \mathcal{P} (X) \), where \(\mathcal{E}\) is not necessarily a \hyperref[def:sigma-algebra]{\(\sigma\)-algebra}. Let \(\left<\mathcal{E}\right>\) be the intersection of all \hyperref[def:sigma-algebra]{\(\sigma\)-algebras} on \(X\) containing \(\mathcal{E}\), then we call \(\left<\mathcal{E} \right>\) the \emph{\hyperref[def:sigma-algebra]{\(\sigma\)-algebra} generated by \(\mathcal{E}\)}.
\end{definition}

\begin{remark}
	Clearly, \(\left<\mathcal{E} \right>\) is the smallest \hyperref[def:sigma-algebra]{\(\sigma\)-algebra} containing \(\mathcal{E}\), and it is unique. To check the uniqueness, we suppose there are two different \(\left<\mathcal{E}\right>_1\) and \(\left<\mathcal{E}\right>_2\) generated from \(\mathcal{E} \). It's easy to show
	\[
		\left<\mathcal{E}\right>_1\subseteq \left<\mathcal{E}\right>_2,
	\]
	and by symmetry, they are equal.
\end{remark}
\begin{eg}
	We see that \(\{\varnothing , B, B^{c} , X\}= \left<\{B\}\right> = \left<\{B^{c} \}\right>\).
\end{eg}

\begin{lemma}\label{lma:lec2-1}
	We have the following.
	\begin{enumerate}[(a)]
		\item Given \(\mathcal{A}\) a \hyperref[def:sigma-algebra]{\(\sigma\)-algebra}, \(\mathcal{E} \subset \mathcal{A} \subset \mathcal{P} (X)\implies \left<\mathcal{E}\right> \subset \mathcal{A} \).
		\item \(\mathcal{E} \subset \mathcal{F} \subset \mathcal{P} (X)\implies \left<\mathcal{E}\right> \subset\left<\mathcal{F} \right>\).
	\end{enumerate}
\end{lemma}
\begin{proof}
	We'll see that after proving the first claim, the second follows smoothly.
	\begin{enumerate}[(a)]
		\item The first claim is trivial, since we know that \(\left<\mathcal{E} \right>\) is the smallest \hyperref[def:sigma-algebra]{\(\sigma\)-algebra} containing \(\mathcal{E} \), then if \(\mathcal{E} \subset \mathcal{A} \), we clearly have \(\left<\mathcal{E} \right>\subset \mathcal{A} \) by the definition.
		\item The second claim is also easy. From the first claim and the definition, we have
		      \[
			      \mathcal{E} \subset \mathcal{F} \subset \left< \mathcal{F} \right> \implies \left< \mathcal{E} \right>\subset \left< \mathcal{F} \right>.
		      \]
	\end{enumerate}
\end{proof}

At this point, we haven't put any specific structure on \(X\). Now we try to describe those spaces with good structure, which will give the space some nice properties.

\begin{definition}[Borel \(\sigma\)-algebra]\label{def:Borel-sigma-algebra}
	For a topological space \(X\), the \emph{Borel \hyperref[def:sigma-algebra]{\(\sigma\)-algebra} on \(X\)}, denotes as \(\mathcal{B}(X)\), is the \hyperref[def:sigma-algebra]{\(\sigma\)-algebra} generated by the collection of all open sets in \(X\).
\end{definition}

\begin{eg}\label{eg:lec2}
	We see that \(\mathcal{B} (\mathbb{R} )\) contains
	\begin{itemize}
		\item \(\mathcal{E}_1 = \left\{(a, b) \mid a < b; a, b\in \mathbb{R} \right\}\).
		\item \(\mathcal{E}_2 = \left\{[a, b] \mid a < b; a, b\in \mathbb{R} \right\}\) since \([a, b] = \bigcap_{n=1}^{\infty} (a - \frac{1}{n}, b + \frac{1}{n})\).
		\item \(\mathcal{E}_3 = \left((a, b]  \mid a < b; a, b\in \mathbb{R} \right)\) since \((a, b] = \bigcap_{n=1}^{\infty} (a, b + \frac{1}{n})\).
		\item \(\mathcal{E}_4 = \left([a, b)  \mid a < b; a, b\in \mathbb{R} \right)\) since \([a, b) = \bigcap_{n=1}^{\infty} (a - \frac{1}{n}, b)\).
		\item \(\mathcal{E}_5 = \left((a, \infty)  \mid a\in \mathbb{R} \right)\) since \((a, \infty) = \bigcup_{n=1}^{\infty} (a , a + n)\).
		\item \(\mathcal{E}_6 = \left([a, \infty)  \mid a\in \mathbb{R} \right)\) since \([a, \infty) = \bigcup_{n=1}^{\infty} [a , a + n)\).
		\item \(\mathcal{E}_7 = \left((-\infty, b) \mid b\in \mathbb{R} \right)\) since \((-\infty, b) = \bigcup_{n=1}^{\infty} (b - n, b)\).
		\item \(\mathcal{E}_8 = \left((-\infty, b] \mid a\in \mathbb{R} \right)\) since \((-\infty, b] = \bigcup_{n=1}^{\infty} (b - n, b]\).
	\end{itemize}
\end{eg}

\begin{proposition}
	\(\mathcal{B} (\mathbb{R} ) = \left<\mathcal{E}_{i}\right> \) for each \(i = 1, \dots, 8\) in the above \hyperref[eg:lec2]{example}.
\end{proposition}
\begin{proof}
	Firstly, we see that \(\mathcal{E} _{i}\subset \mathcal{B} (\mathbb{R} ) \implies \left<\mathcal{E} _{i}\right>\subset \mathcal{B} (\mathbb{R} )\) by \autoref{lma:lec2-1}. Secondly, by \hyperref[def:Borel-sigma-algebra]{definition}, \(\mathcal{B} (\mathbb{R} ) = \left<\mathcal{E} \right>\) where
	\[
		\mathcal{E} = \left\{O\subseteq \mathbb{R}\mid O \text{ is open in \(\mathbb{R}\) } \right\}.
	\]
	It's enough to show \(\mathcal{E} \subset \left<\mathcal{E} _{i}\right>\) since if so, \(\left<\mathcal{E} \right>\subseteq \left<\mathcal{E} _{i}\right>\), and clearly \(\left< \mathcal{E}  \right> \supseteq \left< \mathcal{E}_{i} \right> = \mathcal{B} (\mathbb{R}) \), then we will have \(\left< \mathcal{E}\right> = \left< \mathcal{E} _{i} \right> \). Let \(O\subset \mathbb{R} \) be an open set, i.e., \(O\in\mathcal{E}\). We claim that every open set in \(\mathbb{R} \) is a countable union of disjoint open intervals.\footnote{See \href{https://math.stackexchange.com/questions/318299/any-open-subset-of-bbb-r-is-a-countable-union-of-disjoint-open-intervals}{here}.} Thus,
	\[
		O = \bigcup_{j=1}^{\infty} I_{j},
	\]
	where \(I_{j}\) open interval with the form of \((a, b), (-\infty, b), (a, \infty), (-\infty, \infty)\). For example, \(\mathcal{E} _1\) is trivially true, and
	\[
		(a, b) = \underbrace{\bigcup_{n=1}^{\infty} \underbrace{\left[a+\frac{1}{n}, b-\frac{1}{n}\right]}_{\in \mathcal{E}_2}}_{\in\left<\mathcal{E} _2\right>}
	\]
	shows the case for \(\mathcal{E} _2\) and
	\[
		(a, \infty) = \bigcup_{k=1}^{\infty} (a, a+k)
	\]
	shows the case for \(\mathcal{E} _5\). It's now straightforward to check open intervals are in \(\left<\mathcal{E} _{i}\right>\) for every \(i\).
\end{proof}

Now, to put a structure on a space, we define the following.
\begin{definition*}
	Let \(X\) be a space.
	\begin{definition}[Measurable space]\label{def:measurable-space}
		A \emph{measurable space} is a tuple of \(X\) and a \hyperref[def:sigma-algebra]{\(\sigma\)-algebra} \(\mathcal{A} \) on \(X\), denoted by \((X, \mathcal{A} )\).
	\end{definition}

	In particular, if the \hyperref[def:sigma-algebra]{\(\sigma\)-algebra} is the \hyperref[def:Borel-sigma-algebra]{Borel \(\sigma \)-algebra} of \(X\), then we give it a special name.
	\begin{definition}[Borel space]\label{def:Borel-space}
		A \emph{Borel space} is a tuple of \(X\) and \(\mathcal{B} (X)\), denoted by \((X, \mathcal{B} (X))\).
	\end{definition}
\end{definition*}

\begin{remark}
	The definition of \hyperref[def:Borel-space]{Borel space} means that \(X\) implicitly has some topological structure.
\end{remark}

\begin{definition}[Measurable set]\label{def:measurable-set}
	Given a \hyperref[def:measurable-space]{measurable space} \((X, \mathcal{A} )\), every \(E\in \mathcal{A} \) is a so-called \emph{\(\mathcal{A}-\)measurable set}.
\end{definition}

\section{Measures}
With the definition of \hyperref[def:measurable-space]{measurable space}, we now can refine our \hyperref[def:measure]{measure} function \(\mu\) as follows.

\begin{definition}[Measure]\label{def:measure}
	Given a \hyperref[def:measurable-space]{measurable space} on \((X, \mathcal{A}) \), a \emph{measure} is a function
	\[
		\mu\colon \mathcal{A} \to [0, \infty]
	\]
	such that
	\begin{itemize}
		\item\label{def:measure-null-empty-set}(null empty set) \(\mu(\varnothing ) = 0\).
		\item\label{def:measure-countable-additivity}(countable additivity) \(\mu\left(\bigcup_{i=1}^{\infty} A_{i}\right) = \sum_{i=1}^{\infty}\mu(A_{i})\) if \(A_1, A_2, \dots \in \mathcal{A}\)
		      are disjoint.
	\end{itemize}
\end{definition}

\begin{definition}[Measure space]\label{def:measure-space}
	We denote \((X, \mathcal{A} , \mu)\) as so-called a \emph{measure space} given \(\mu\) is the \hyperref[def:measure]{measure} on \((X, \mathcal{A} )\).
\end{definition}

\begin{notation}
	We denote \([0, \infty] \coloneqq [0, \infty) \cup \{\infty\}\).
\end{notation}

\begin{remark}
	We only want \hyperref[def:measure-countable-additivity]{countable additivity} but not uncountable additivity.
\end{remark}
\begin{explanation}
	Consider the most intuitive \hyperref[def:measure]{measure} on \((\mathbb{R} , \mathcal{B} (\mathbb{R}))\). Since we have
	\[
		(0, 1] = (1 / 2, 1]\cup (1 / 4, 1 / 2]\cup (1 / 8, 1 / 4]\cup \dots
	\]
	and also
	\[
		(0, 1] = \bigcup_{x\in (0, 1]}\{x\}.
	\]
	In the first case, we are claiming that
	\[
		1
		= \frac{1}{2} + \frac{1}{4} + \frac{1}{8} + \dots
		= \mu((1 / 2, 1]) + \mu((1 / 4, 1 / 2]) + \mu((1 / 8, 1 / 4]) + \dots ;
	\]
	while in the second case, we are claiming that
	\[
		1 = \sum_{x\in(0, 1]} 0
	\]
	since \(\mu(\{ x \} ) = 0\) for \(x\in \mathbb{R} \), which is clearly not what we want.
\end{explanation}

\begin{eg}[Counting measure]\label{eg:counting-measure}
	For any \((X, \mathcal{A})\), we let \(\mu(A) \coloneqq \vert A \vert\). This is the so-called \emph{counting \hyperref[def:measure]{measure}}.
\end{eg}
\begin{eg}[Dirac-Delta measure]\label{eg:Dirac-Delta measure}
	Let \(x_0\in X\). For any \((X, \mathcal{A})\), the \emph{Dirac-Delta \hyperref[def:measure]{measure} at \(x_0\)} is
	\[
		\mu(A) = \begin{dcases}
			1, & \text{ if } x_0\in A;   \\
			0, & \text{ if } x_0\notin A
		\end{dcases}
	\]
	for every \(A\in \mathcal{A} \).
\end{eg}
\begin{eg}
	For \((\mathbb{N}, \mathcal{P} (\mathbb{N} ))\), given \(A\in \mathcal{P} (\mathbb{N} )\),
	\[
		\mu(A) = \sum_{i\in A}a_{i}
	\]
	where \(a_1, a_2, \dots \in [0, \infty) \).
\end{eg}