\chapter{Introduction}
\lecture{1}{30 Aug. 14:30}{Introduction}

We first briefly review different kinds of vector spaces.

\section{Linear Space}

\begin{definition}[Linear vector space]\label{def:linear-vector-space}
	A set with operations of addition and multiplication (by a scalar) is called a \emph{linear vector space}.
\end{definition}

\begin{eg}
	Denote the multiplicative scalar by \(\lambda \), then
	\begin{itemize}
		\item \(\lambda \in \mathbb{\MakeUppercase{r}} \implies\) real vector space.
		\item \(\lambda \in \mathbb{\MakeUppercase{c}} \implies \) complex vector space
	\end{itemize}
\end{eg}

\begin{lemma}
	Given \(E\) a \hyperref[def:linear-vector-space]{linear vector space}, if \(v, w\in E\), \(\lambda , \mu \in \mathbb{\MakeUppercase{r}} \) (or \(\mathbb{\MakeUppercase{c}} \)), then \(\lambda v + \mu w\in E\).
\end{lemma}

we also have usual rules of associativity and commutativity.

\begin{eg}
	\(\mathbb{\MakeUppercase{r}} ^n\) a \(n\) dimensional \hyperref[def:linear-vector-space]{linear vector space}, \(\mathbb{\MakeUppercase{c}} ^n\) a \(n\) dimensional complex \hyperref[def:linear-vector-space]{linear vector space}.
\end{eg}

We concentrate on \(\infty \) dimensional \hyperref[def:linear-vector-space]{linear vector space}.

\begin{eg}
	Let \(K\) is a compact Hausdorff space, then
	\[
		E = \left\{ f\colon K\to \mathbb{\MakeUppercase{r}} \mid f(\cdot) \text{ is continuous}  \right\}.
	\]
	We then see that \(E\) is an \(\infty\)	dimensional \textbf{real} \hyperref[def:linear-vector-space]{linear vector space}.
\end{eg}

\section{Quotient Space}
Observe that a \hyperref[def:linear-vector-space]{linear vector space} can have many subspaces. Say \(E\) is a \hyperref[def:linear-vector-space]{linear vector space}, and \(E_1 \subset E\) where \(E_1\) is a proper subspace, i.e., \(E_1 \neq E\).

\begin{definition}[Quotient Space]\label{def:quotient-space}
	The \emph{quotient space} \(\quotient{E}{E_1} \) is the set of equivalence classes of vectors in \(E\) where equivalence is given by \(x\sim y\) if \(x - y\in E_1\). Additionally, denote \([x]\) as the equivalence class of \(x\in E\), i.e., \([x] = x + E_1\).
\end{definition}

Note that \(\quotient{E}{E_1} \) is a \hyperref[def:linear-vector-space]{linear vector space} since if \(x_1 + x_2\in E\), \([x_1] + [x_2] = [x_1 + x_2]\), and also, \(\lambda [x] = [\lambda x]\) for \(\lambda \in \mathbb{\MakeUppercase{r}} \) or \(\mathbb{\MakeUppercase{c}} \), i.e., \(v, w\in \quotient{E}{E_1} \), \(\lambda , \mu \in\mathbb{\MakeUppercase{r}} \) or \(\mathbb{\MakeUppercase{c}} \) implies \(\lambda v + \mu w\in E\).

\begin{definition}[Codimension]\label{def:codimension}
	If \(\quotient{E}{E_1} \) has finite dimension, then the dimension of \(\quotient{E}{E_1} \) is called the \emph{cdimension} of \(E_1\) in \(E\).
\end{definition}

\begin{eg}
	There exists the case that \(\dim(E) = \infty \), \(\dim(E_1) < \infty\) where \(\dim(\quotient{E}{E_1} )<\infty \).
\end{eg}
\begin{explanation}
	Let \(E = \left\{ f\colon K\to \mathbb{\MakeUppercase{r}} \mid f(\cdot) \text{ continuous}  \right\} \), and \(E_1 = \left\{ f\in E\colon f(k_1) = 0 \right\} \) where \(k_1\in K\) is fixed. We see that the dimension of \(\quotient{E}{E_1} \) is exactly \(1\) since \(\quotient{E}{E_1} \) is the set of constant functions.
\end{explanation}

\begin{theorem}
	If \(E\) is finite dimensional, then \(\mathop{\mathrm{codim}}(E_1) + \dim (E_1) = \dim(E)\)
\end{theorem}

\begin{definition}[Linear operator]\label{def:linear-operator}
	A map \(T\colon E\to F\) between 2 \hyperref[def:linear-vector-space]{linear spaces} is a \emph{linear operator} if it preserves the properties of addition and multiplication by a scalar, i.e., \(T(\lambda v + \mu w) = \lambda T(v) + \mu T(w)\) for \(v, w\in E\) and \(\lambda , \mu \in \mathbb{\MakeUppercase{r}}\) or \(\mathbb{\MakeUppercase{c}}\).
\end{definition}

\begin{definition*}
	Given a \hyperref[def:linear-operator]{inear operator} \(T\colon E \to F\) we have the following.
	\begin{definition}[Kernel]
		The \emph{kernel} of \(T\) is the subspace \(\ker(T) = \left\{ x\in E\mid Tx=0 \right\} \).
	\end{definition}

	\begin{definition}[Image]
		The \emph{image} of \(T\) is the subspace \(\im(T) = \left\{ Tx\in F\mid x\in E \right\} \).
	\end{definition}
\end{definition*}

\section{Normed Spaces}
We review some basic notions.

\begin{definition}[Norm]\label{def:norm}
	Let \(E\) be a \hyperref[def:linear-vector-space]{linear vector space}. A \emph{norm} \(\left\lVert \cdot \right\rVert \colon E \to \mathbb{\MakeUppercase{r}} \) on \(E\) is a function from \(E\) to \(\mathbb{\MakeUppercase{r}} \) with the properties:
	\begin{enumerate}[(a)]
		\item \(\left\lVert x\right\rVert \geq 0\) and \(\left\lVert x\right\rVert =0 \iff x=0\).
		\item \(\left\lVert \lambda x\right\rVert = \left\vert \lambda  \right\vert \left\lVert x\right\rVert\), \(\lambda \in\mathbb{\MakeUppercase{r}} \) or \(\mathbb{\MakeUppercase{c}} \).
		\item \(\left\lVert x+y\right\rVert \leq \left\lVert x\right\rVert + \left\lVert y\right\rVert \).
	\end{enumerate}
\end{definition}

\begin{notation}[Dilation]
	We say that the second condition is the \emph{dilation} property.
\end{notation}

\begin{definition}[Normed vector space]\label{def:normed-vector-space}
	A \hyperref[def:linear-vector-space]{linear vector space} \(E\) equipped with a \hyperref[def:norm]{norm} \(\left\lVert \cdot\right\rVert \) is called a \emph{normed vector space}.
\end{definition}

\begin{remark}[Induced metric space]
	A \hyperref[def:normed-vector-space]{normed vector space} \(E\) induces a \emph{metric space} with metric \(d(x, y) = \left\lVert x- y\right\rVert \), where the metric has properties
	\begin{enumerate}[(a)]
		\item \(d(x, y) \geq 0\). Also, \(d(x, x) = 0\) and \(d(x, y)\) implies \(x =y\).
		\item \(d(x, y) = d(y, x)\).
		\item \(d(x, z) \leq d(x, y) + d(y, z)\).
	\end{enumerate}
\end{remark}

\begin{eg}[Bounded sequences \(\ell_\infty\)]
	Let \(\ell _\infty \) be the space of bounded sequences \(x = (x_1, x_2, \ldots )\) with \(x_i\in \mathbb{\MakeUppercase{r}} \) for \(i = 1, 2, \ldots \). Then we define \(\left\lVert x\right\rVert = \left\lVert x\right\rVert _\infty = \sup _{i \geq 1}\left\vert x_i \right\vert \).
\end{eg}

\begin{eg}[Absolutely summable sequences \(\ell _1\)]
	Let \(\ell _1\) be the space of absolutely summable sequences \(x = (x_1, x_2, \ldots)\) and \(\sum_{i=1}^{\infty} \left\vert x_i \right\vert < \infty\). Then we define \(\left\lVert x\right\rVert = \left\lVert x\right\rVert _1 = \sum_{i=1}^{\infty} \left\vert x_i \right\vert < \infty\).
\end{eg}

\begin{eg}[Continuous functions \(C(k)\)]
	The space \(C(k)\) of continuous functions \(f\colon K\to \mathbb{\MakeUppercase{r}} \) where \(K\) is compact Hausdorff. Then we define \(\left\lVert f\right\rVert = \left\lVert f\right\rVert _\infty = \sup _{x\in K}\left\vert f(x) \right\vert \).
\end{eg}

\subsection{Geometry of Normed Spaces}
\begin{definition}[Ball]\label{def:ball}
	A (closed) \emph{ball} centered at a point \(x_0\in E\) with radius \(r>0\) is the set \(B(x_0, r) = \left\{ x\in E\mid \left\lVert x - x_0\right\rVert \leq r \right\} \).
\end{definition}

\begin{definition}[Sphere]\label{def:sphere}
	The \emph{sphere} centered at \(x_0\) with radius \(r>0\) is the set \(S(x_0, r) = \left\{ x\in E\mid \left\lVert x - x_0\right\rVert = r\right\} \).
\end{definition}

\begin{remark}
	We see that \(S(x_0, r)\) is the \textbf{boundary} of \(B(x_0, r)\), i.e., \(S(x_0, r) = \partial B(x_0, r)\).
\end{remark}

\begin{note}[Nonequivalency in infinite dimensional spaces]
	We know that in finite dimensional, all \hyperref[def:norm]{norms} are equivalent, which is not true for infinite dimensional vector spaces.

	This has something to do with the geometry of \hyperref[def:ball]{balls}.
\end{note}

Explicitly, \hyperref[def:ball]{balls} can have different geometries depending on the properties of the \hyperref[def:norm]{norms}. We see that an \(\left\lVert \cdot\right\rVert _{\infty}\) can have multiple supporting hyperplane at the corner, while for an \(\left\lVert \cdot\right\rVert _2\) can have only one at each point.

Also, unit \hyperref[def:ball]{balls} for \(\left\lVert \cdot\right\rVert _1\) is also a \textbf{square}, where we have
\[
	B(0, 1) = \left\{ x = (x_1, x_2, \ldots)\mid -1 < y_{\epsilon } < 1 \forall \epsilon\right\}
\]
such that \(y_{\epsilon } = \sum_{i=1}^{\infty} \epsilon _i x_i \), \(\epsilon _i = \pm 1\) and \(\epsilon = (\epsilon _1, \epsilon _2, \ldots  )\).

We see that different \hyperref[def:norm]{norms} give different geometry, but they have important common features, most notably, convexity properties.

\begin{definition}[Convex set]\label{def:convex-set}
	Given \(E\) a \hyperref[def:linear-vector-space]{linear vector space}, a set \(K\subset E\) is \emph{convex} if \(x, y\in K\) and \(0 \leq \lambda \leq 1\), we have \(\lambda x e (1 - \lambda )y\in K\).
\end{definition}

\begin{definition}[Convex function]\label{def:convex-function}
	Given \(E\) a \hyperref[def:linear-vector-space]{linear vector space}, a function \(f\colon E\to \mathbb{\MakeUppercase{r}} \) is called \emph{convex} if
	\[
		f(\lambda x + (1 - \lambda )y) \leq \lambda f(x) + (1 - \lambda )f(y)
	\]
	for \(x, y\in E\), \(0 \leq \lambda \leq 1\).
\end{definition}

\begin{remark}
	If \(f\colon E \to \mathbb{\MakeUppercase{r}} \) is a \hyperref[def:convex-function]{convex function}, then for any \(M\in \mathbb{\MakeUppercase{r}} \) the set \(\left\{ x\in E\mid f(x) \leq M \right\} \) is \hyperref[def:convex-set]{convex}.
\end{remark}

The upshot is that \hyperref[def:norm]{norms} are \hyperref[def:convex-function]{convex}, and the unit \hyperref[def:ball]{balls} are \hyperref[def:convex-set]{convex} as well.
