\lecture{4}{17 Jan. 14:30}{Theory and Axioms}
\begin{eg}
	Let \(\mathcal{L} _1 = \left\{ E \right\} \), where \(E\) is a binary relation; and \(\mathcal{L} _2 = \left\{ V, E, I \right\} \), where \(V, E\) are unary relations and \(I\) is a binary relation such that \(I(v, e)\) for \(v\in V\), \(e\in E\) means that \(v\) is a vertex on edge \(e\).
\end{eg}

Let \(G\) be a graph, viewed as a \(\mathcal{L} _1\)-structure. A substructure of \(G\) is an induced subgraph \(H \subseteq G\) such that any edge in \(G\) between two vertices of \(H\) is in \(H\). If we view \(G\) as an \(\mathcal{L} _2\)-substructure, a substructure is a subgraph \(H\) such that \(H\) has some vertices and edges from \(G\).\footnote{But there might be edges in \(H\) with no vertices, which can be fixed by having two functions \(I_1(e)=v\), \(I_2(e)=w\) when \(e\colon v \to w\).}

For example, \(vEw\) is \hyperref[not:quantifier-free-formula]{quantifier-free} in \(\mathcal{L} _1\), while
\[
	\exists \ (v\in V \land w\in V \land e\in E \land I(v, e) \land I(w, e))
\]
is not \hyperref[not:quantifier-free-formula]{quantifier-free} in \(\mathcal{L} _2\).

\begin{definition}[Theory]\label{def:theory}
	An \emph{\(\mathcal{L}\)-theory} is a set of \(\mathcal{L} \)-sentences.
\end{definition}

\begin{definition}[Model]\label{def:model}
	We say that \(\mathcal{M} \) is a \emph{model} of a \hyperref[def:theory]{theory} \(T\), written as \(\mathcal{M} \models T\), if \(\mathcal{M} \models \varphi \) for all \(\varphi \in T\).
\end{definition}

\begin{note}
	Not every \hyperref[def:theory]{theory} has a \hyperref[def:model]{model}, e.g., \(\left\{ \forall x\ x \neq x \right\} \).
\end{note}

\begin{definition}[Satisfiable]\label{def:satisfiable}
	If a \hyperref[def:theory]{theory} has a \hyperref[def:model]{model}, we say that it is \emph{satisfiable}.
\end{definition}

\begin{definition}[Elementary class]\label{def:elementary-class}
	A class \(\mathcal{K} \) of \(\mathcal{L} \) -structures is called an \emph{elementary class} if there is an \(\mathcal{L} \)-theory \(T\) such that
	\[
		\mathcal{K} = \left\{ \mathcal{M} \mid \mathcal{M} \models T \right\}.
	\]
\end{definition}

Start with an \(\mathcal{L} \)-structure \(\mathcal{M} \), then the theory of \(\mathcal{M} \) is
\[
	\mathop{\mathrm{Th}}(\mathcal{M} ) = \left\{ \varphi \mid \mathcal{M} \models \varphi  \right\}.
\]
It's easy to see that \(\mathcal{M} \models \mathop{\mathrm{Th}}(\mathcal{M} ) \).

\begin{remark}
	\(\mathop{\mathrm{Th}}(\mathcal{M} ) \) characterizes the structures satisfying the same sentences as \(\mathcal{M} \).
\end{remark}


\begin{definition}[Complete]\label{def:complete}
	A \hyperref[def:theory]{theory} \(T\) is \emph{complete} if for any sentence \(\varphi \), either \(\varphi \in T\) or \(\lnot \varphi \in T\).
\end{definition}

\begin{remark}
	\(\mathop{\mathrm{Th}}(\mathcal{M} ) \) is \hyperref[def:complete]{complete}.
\end{remark}

\begin{definition}[Elementarily equivalent]\label{def:elementarily-equivalent}
	\(\mathcal{M} \) and \(\mathcal{N} \) are \emph{elementarily equivalent}\(\mathcal{M} \equiv \mathcal{N} \) if for all sentences \(\varphi \),
	\[
		\mathcal{M} \models \varphi \iff \mathcal{N} \models \varphi .
	\]
\end{definition}

\begin{remark}[Non-standard model of arithmetic]
	There are \(\mathcal{N} \models \mathop{\mathrm{Th}}(\mathbb{N})\), but \(\mathcal{N} \) is not isomorphic to \(\mathbb{N} \). \(\mathcal{N} \) is called a \emph{non-standard model of arithmetic}. \(\mathcal{N} \) has \emph{infinite element} larger than all of \(\mathbb{M} \). Here, \(\mathbb{N} =(\mathbb{N} , 0, 1, +, \cdot, - )\)
\end{remark}

\begin{eg}
	\(\mathbb{Z} \oplus \mathbb{Z} \not\equiv \mathbb{Z} \) as groups.
\end{eg}
\begin{explanation}
	We will show this on the homework.
\end{explanation}

The other way to define a \hyperref[def:theory]{theory} is to write down axioms.

\begin{eg}[Infinite set]
	Let \(\mathcal{L} = \varnothing \), and let \(T\) consist of
	\[
		\varphi _n \coloneqq \exists x_1 \ldots \exists x_n\ \bigwedge _{i \neq j} x_i \neq x_j.
	\]
\end{eg}

\begin{eg}[Linear order]
	Let \(\mathcal{L} =\left\{ \leq \right\} \), and let \(T\) consist of the axioms of linear orders, e.g.,
	\[
		\forall x\forall y\ (x \leq y \land y \leq x \to x=y).
	\]
\end{eg}

There are other interesting \hyperref[def:theory]{theories} of linear orders, e.g., dense ones.

\begin{eg}[Dense linear order]
	Consider
	\[
		\forall x \forall y\ (x < y \to  \exists z\ x < z < y),
	\]
	where we use \(a < b\) as shorthand of saying \(a \leq b \land a \neq b\).
\end{eg}

\begin{eg}[Group]
	In \(\mathcal{L} _{\text{group} } = \left\{ e, \cdot, ^{-1}  \right\} \), let \(T\) be the group axioms.
\end{eg}

Other theories of groups include Abelson group, divisible, etc.

\begin{definition}[Finitely axiomatizable]\label{def:finitely-axiomatizable}
	A \hyperref[def:theory]{theory} is \emph{finitely axiomatizable} if it has a finite axiomatization.
\end{definition}

Given a \hyperref[def:theory]{theory}, consider \(T^{\models } = \left\{ \varphi \colon T \models \varphi \right\}\),\footnote{Recall \autoref{def:logical-consequence}.} so \(\mathcal{M} \models T\) if and only if \(\mathcal{M} \models T^{\models }\). Often we think of \(T\) and \(T^{\models }\) as the same. A \hyperref[def:theory]{theory} \(T\) is \hyperref[def:finitely-axiomatizable]{finitely axiomatizable} if there is a finite \(\Phi \) such that \(T^{\models } = \Phi ^{\models }\).

\begin{definition}[Elementary embedding]\label{def:elementary-embedding}
	Let \(\mathcal{M} \) and \(\mathcal{N} \) be \(\mathcal{L} \)-structures, and \(f\colon \mathcal{M} \to \mathcal{N} \) an \(\mathcal{L} \)-embedding. Then \(f\) is an \emph{elementary embedding} if for any formula \(\varphi (\overline{x} )\) and \(\overline{a} \in M\),
	\[
		\mathcal{M} \models \varphi (\overline{a} )\iff \mathcal{N} \models \varphi (f(\overline{a} )).
	\]
\end{definition}

\begin{eg}
	As groups, \(\mathbb{Z} \hookrightarrow \mathbb{Q} \) is not \hyperref[def:elementary-embedding]{elementary}. In fact, \(\mathbb{Z} \not\equiv \mathbb{Q} \). Wheres, if \(f\colon \mathcal{M} \hookrightarrow \mathcal{N} \) is an \hyperref[def:elementary-embedding]{elementary embedding}, \(\mathcal{M} \equiv \mathcal{N} \).\footnote{And also much more is true.}
\end{eg}

\begin{proposition}
	Every \hyperref[def:isomorphism]{isomorphism} is an \hyperref[def:elementary-embedding]{elementary embedding}.
\end{proposition}
\begin{proof}
	Let \(f\colon \mathcal{M} \to  \mathcal{N} \) be an \hyperref[def:isomorphism]{isomorphism}. We will argue by induction on \hyperref[def:formula]{formulas} \(\varphi \), that for all \(\overline{a} \in M\),
	\[
		\mathcal{M} \models \varphi (\overline{a} ) \iff \mathcal{N} \models \varphi (f(\overline{a} )).
	\]
	Firstly, observe that all cases except quantifiers are the same as \autoref{prop:lec3-1}. For quantifiers, suppose that \(\varphi (\overline{x} )\) is \(\exists y\ \psi (\overline{x} , y)\) and \(\mathcal{M} \models \varphi (\overline{a} )\). This means that there is \(b\in M\) such that \(\mathcal{M} \models \psi (\overline{a} , b)\). By the induction hypothesis, \(\mathcal{N} \models \psi (f(\overline{a} , f(b)))\), so \(\mathcal{N} \models \varphi (f(\overline{a} ))\).

	Now suppose \(\mathcal{N} \models \varphi (f(\overline{a} ))\), then there is \(c\in N\) such that \(\mathcal{N} \models \psi (f(\overline{a} ), c)\). Since \(f\) is an \hyperref[def:isomorphism]{isomorphism}, so there is a \(b\in M\) such that \(f(b) = c\). By the induction hypothesis, \(\mathcal{M} \models \psi (\overline{a} , b)\), so \(\mathcal{M} \models \varphi (\overline{a} )\).
\end{proof}

\begin{corollary}
	If \(\mathcal{M} \cong \mathcal{N} \), then \(\mathcal{M} \equiv \mathcal{N} \).
\end{corollary}

\begin{definition}[Definable set]\label{def:definable-set}
	Let \(\mathcal{M} \) be an \(\mathcal{L} \)-structure, and let \(X \subseteq M ^n\). \(X\) is \emph{definable} if there is a \hyperref[def:formula]{formula} \(\varphi (x_1, \ldots , x_n, \overline{y} )\) and \(\overline{b} \in M\) such that
	\[
		X = \left\{ \overline{a} \in M^n \mid \mathcal{M} \models \varphi (\overline{a} , \overline{b} ) \right\}.
	\]
\end{definition}

\begin{notation}[Define]\label{not:define}
	We say that \(\varphi (\overline{x} , \overline{b} )\) \emph{defines} \(X\) over \(\overline{b} \), written as \(X = \varphi (\mathcal{M} , \overline{b} )\).
\end{notation}

\begin{notation}[Parameter]\label{not:parameter}
	The tuple \(\overline{b} \) is called the \emph{parameters} when \(X\) is \hyperref[def:definable-set]{definable} over \(\overline{b} \).
\end{notation}

\begin{remark}
	Sometimes \(X\) is \hyperref[def:definable-set]{definable} without \hyperref[not:parameter]{parameters}, or \hyperref[def:definable-set]{definable} over \(\varnothing \).
\end{remark}

\begin{eg}
	Take \(\mathbb{R} =(\mathbb{R} , 0, 1, +, \cdot, -)\) in \(\mathcal{L} _{\text{ring} }\), then
	\[
		\leq = \left\{ (a, b) \colon a \leq b\right\}
	\]
	is \hyperref[def:definable-set]{definable}.
\end{eg}

\begin{eg}[Langrange's four-square theorem]
	Let \(\mathbb{Z} =(\mathbb{Z} , +, -, \cdot, 0, 1)\). Then \(\mathbb{N} \) is \hyperref[def:definable-set]{\(\varnothing \)-definable} in \(\mathbb{Z} \) by
	\[
		\mathbb{N} = \left\{ z\in \mathbb{Z} \colon \exists u, v, x, y\ u^2 + v^2 + x^2 + y^2 = z \right\}.
	\]
	This follows from the \href{https://en.wikipedia.org/wiki/Lagrange%27s_four-square_theorem}{Langrange's four-square theorem}, which says that every natural number is the sum of four squares.
\end{eg}

\begin{eg}
	\(\mathbb{Z} \) is \hyperref[def:definable-set]{\(\varnothing \)-definable} in \(\mathbb{Q} = (\mathbb{Q} , +, -, \cdot, 0, 1)\). This is a result of Julia Robinson, and the formula is very complicated.
\end{eg}

\begin{problem*}
	How does one show that a set is not \hyperref[def:definable-set]{definable}? For example, \(\mathbb{R} \) are not \hyperref[def:definable-set]{definable} in \(\mathbb{C} = (\mathbb{C} , 0, 1, +, \cdot, -)\).
\end{problem*}