\lecture{8}{26 Sep.\ 10:30}{Local Search for \(k\)-Median}
We'll now see a completely different algorithm which solve \autoref{prb:k-median} with \((3 + \epsilon )\)-approximation ratio by local search.
\subsection{Local Search}
The idea is to iteratively improve the current solution. We first see the algorithm.

\begin{algorithm}[H]\label{algo:k-median-local-search}
	\DontPrintSemicolon
	\caption{\hyperref[prb:k-median]{\(k\)-Median} -- Local Search}
	\KwData{A set of clients \(P\subseteq X\), a set of (possible) facilities \(Q\subseteq X\), \(k\in \mathbb{N} \), width \(w\)}
	\KwResult{A set of opened facilities \(Q^\prime \subseteq Q\) with \(\left\vert Q ^\prime \right\vert = k \)}
	\BlankLine

	\(Q^\prime \gets \) arbitrary \(k\) centers in \(Q\)\;
	\While(\label{algo:k-median-local-search-while}){\(\exists Q^{\prime\prime}\) s.t. \(\left\vert Q^{\prime\prime} \right\vert = k \) and \(\mathop{\mathrm{cost}}(Q^{\prime\prime}) < \mathop{\mathrm{cost}}(Q^\prime )\) and \(\left\vert Q^\prime \triangle Q^{\prime\prime} \right\vert \leq w\)\footnote{The symmetric difference \(A \triangle B\) is defined as \(A \triangle B \coloneqq (A \setminus B) \cup (B\setminus A)\).}}{
		\(Q^\prime \gets Q^{\prime\prime}\)\;
	}
	\Return{\(Q^\prime \)}\;
\end{algorithm}

\begin{remark}[Runtime]
	In \autoref{algo:k-median-local-search-while}, each iteration in \autoref{algo:k-median-local-search} takes \((n+m)^{O(w)}\) time for \(n \coloneqq \left\vert P \right\vert \) and \(m \coloneqq \left\vert Q \right\vert \). But we have no control of how many iterations \autoref{algo:k-median-local-search} might take since we might decrease the cost by a little each time hence we might fall into exponentially many updates. To solve this, we can ask for
	\[
		\mathop{\mathrm{cost}}(Q^{\prime\prime}) < \mathcolor{red}{(1 - \epsilon )}\mathop{\mathrm{cost}}(Q^\prime )
	\]
	instead to make sure we decrease a reasonable amount each time, which guarantees that we can bound the number of iterations by
	\[
		\log_\frac{1}{1-\epsilon } \left( \frac{\mathop{\mathrm{cost}}(\text{starting } Q^\prime )}{\OPT} \right).
	\]
\end{remark}

To do the analysis, first note that for any solution \(Q^\prime \) output from \autoref{algo:k-median-local-search}, we have that there exists no \(Q^{\prime\prime}\) such that \(\left\vert Q^\prime \triangle Q^{\prime\prime} \right\vert \leq w\), \(\left\vert Q^{\prime\prime} \right\vert = k\) and \(\mathop{\mathrm{cost}}(Q^{\prime\prime}) < \mathop{\mathrm{cost}}(Q^\prime )\).

\begin{note}[Local optimum]
	We say this \(Q^\prime \) is a \emph{local optimum}.
\end{note}

Let \(Q^{\ast} \subseteq Q\) be the optimal solution, and without loss of generality (by duplicating facilities), assume \(Q^\prime \cap Q^{\ast} = \varnothing \). We define something called \hyperref[not:swap]{swap}.

\begin{notation}[Swap]\label{not:swap}
	A \emph{swap} \(S \subseteq Q^\prime \cup Q^{\ast} \) satisfies \(\left\vert S \cap Q^\prime  \right\vert = \left\vert S \cap Q^{\ast}  \right\vert \leq w / 2\).
\end{notation}

\begin{note}
	From local optimality of \(Q^\prime \), for any \hyperref[not:swap]{swap} \(S\), \(\mathop{\mathrm{cost}}(Q^\prime ) \leq \mathop{\mathrm{cost}}(Q^\prime \triangle S)\).
\end{note}

Now, consider constructing \hyperref[not:swap]{swaps} \(S_1, \dots, S_t\) with weights \(p_1, \dots , p_t \in \mathbb{R} ^+\) such that \(\mathop{\mathrm{cost}}(Q^\prime ) \leq \mathop{\mathrm{cost}}(Q^\prime \triangle S_i)\) for all \(i\), we have
\begin{equation}\label{eq:lec8}
	\sum_{i=1} ^t p_i \cdot (\mathop{\mathrm{cost}}(Q^\prime ) - \mathop{\mathrm{cost}}(Q^\prime \triangle S_i)) \leq 0.
\end{equation}
Our goal is to show that \autoref{eq:lec8} implies \(\mathop{\mathrm{cost}}(Q^\prime ) \leq \alpha\cdot \mathop{\mathrm{cost}}(Q^{\ast} )\) for some \(\alpha \in \mathbb{R} ^+\). To do this, we require the set of \hyperref[not:swap]{swaps} to have the following properties.

\begin{enumerate}[(a)]
	\item For all \(j\in Q^{\ast} \), \(\sum_{S_i \ni j}p_i = 1 \), and let \(p^\prime \coloneqq \max _{j\in Q^\prime }\sum_{S_i\ni j}p_i\).
	\item For all \(j\in Q^{\ast} \), let \(\pi (j)\in Q^\prime \) be the facility closest to \(j\). Then if \(S_i\) contains \(j\in Q^\prime \), \(\pi ^{-1} (j)\subseteq S_i\).\footnote{In particular, if \(\left\vert \pi ^{-1} (j) \right\vert > w / 2 \), then \(j\) is not contained in any \hyperref[not:swap]{swap}.}
\end{enumerate}

The existence of such \hyperref[not:swap]{swaps} family is ensured by the following lemma.
\begin{lemma}\label{lma:lec8}
	There exists a family of \hyperref[not:swap]{swaps} \(S_1, \dots, S_t\) with weights \(p_1, \dots  , p_t\) such that \(\forall j\in Q^{\ast} \), \(\sum_{S_i \ni j}p_i = 1\) and if \(j\in S_i\), \(\pi ^{-1} (j)\subseteq S_i\) with \(p^\prime = \max _{j\in Q^\prime }\sum_{S_i\ni j}p_i = 1 + 2 / w\).
\end{lemma}
\begin{proof}
	For all \(j\in Q^\prime \), we call \(j\)
	\begin{itemize}
		\item \emph{big}: if \(\left\vert \pi ^{-1} (j) \right\vert > w / 2\).
		\item \emph{small}: if \(\left\vert \pi ^{-1} (j) \right\vert \in [1, w / 2]\).
		\item \emph{lonely}: if \(\left\vert \pi ^{-1} (j) \right\vert = 0\).
	\end{itemize}
	Then for each small or big \(j\), we create a \emph{group} \(G_j\) that contains \(\pi ^{-1} (j)\), \(j\) and \(\left\vert \pi ^{-1} (j) \right\vert - 1\) lonely facilities (denote as \(R_j \subseteq Q^\prime \)). We see that \(\left\vert G_j \right\vert = 2 \left\vert \pi ^{-1} (j) \right\vert \), and we can ensure each lonely facility belongs to exactly \(1\) group, i.e., \(\exists G_1, \dots G_r\) such that each facility belongs to exactly \(1\) group. It's now clear that how we should create \hyperref[not:swap]{swaps} and their corresponding weight:
	\begin{enumerate}[(a)]
		\item For small \(j\), let \(G_j\) be a \hyperref[not:swap]{swap} with weight \(1\).
		\item For big \(j\), let \(w^\prime \coloneqq \left\vert \pi ^{-1} (j) \right\vert \), then for any \(S \subseteq \pi ^{-1} (j)\) and \(T \subseteq R_j\) with \(\left\vert S \right\vert = \left\vert T \right\vert = w / 2\), we let \((S \cup T)\) be a \hyperref[not:swap]{swap} with weight \(1 / \left( \binom{w^\prime -1}{w / 2 - 1} \cdot \binom{w^\prime - 1}{w / 2} \right) \).\footnote{Notice that since \(j\) is big, so \(j\) can't be in any \hyperref[not:swap]{swap}, so we have only \(w^\prime - 1\) to choose from.}
	\end{enumerate}
	Since for every \(j^{\ast} \in Q^{\ast} \), there is only one group containing \(j^{\ast}\), to verify \(\sum_{S_i \ni j^{\ast} }p_i = 1 \), we see that
	\begin{enumerate}[(a)]
		\item \(j^{\ast} \) is containing in \(G_j\) for \(j\) small: In this case, we have one \hyperref[not:swap]{swap}, i.e., \(G_j\) itself with weight \(1\).
		\item \(j^{\ast} \) is containing in \(G_j\) for \(j\) big: In this case, since every such \hyperref[not:swap]{swap} created inside \(G_j\) contains \(j^{\ast} \) and has uniform weight, it sums up to \(1\).
	\end{enumerate}

	Finally, we want to show that \(p^\prime = \max _{j\in Q^\prime }\sum_{S_i \ni j}p_i = 1 + 2 / w \). But this is also easy since given \(j\in Q^\prime \), the summation is inside \(G_j\), and in particular, \(S_i\) is inside \(G_j\) as well. Then
	\begin{enumerate}[(a)]
		\item \(j\) is small: Only swap is \(G_j\) itself with weight \(1\).
		\item \(j\) is big: \(j\) can't even be in one \hyperref[not:swap]{swap}, hence the sum is \(0\).
		\item \(j\) is lonely: In this case, we have
		      \[
			      \sum_{S_i \ni j} p_i = \frac{1}{\binom{w^\prime - 1}{w / 2 - 1} \binom{w^\prime - 1}{w / 2}}\cdot \binom{w^\prime }{w / 2}\binom{w^\prime -2}{w / 2 - 1} = \frac{w^\prime }{w / 2} \frac{w/2}{w^\prime -1} = \frac{w^\prime }{w^\prime -1} \leq 1 + \frac{2}{w}.
		      \]
	\end{enumerate}
	Taking the maximum, we have \(p^\prime = 1 + 2 / w\) as desired.
\end{proof}

With \autoref{lma:lec8}, we're ready to prove the following.

\begin{theorem}\label{thm:lec8}
	\autoref{algo:k-median-local-search} is a \((3 + \epsilon)\)-approximation algorithm for arbitrary small \(\epsilon > 0\).
\end{theorem}
\begin{proof}
	Fix \(i\in P\), we analyze how it contributes to the left-hand side of \autoref{eq:lec8}. Let \(j^\prime \in Q^\prime \) and \(j^{\ast} \in Q^{\ast} \) be facilities closest to \(i\), and \(d_i^\prime \coloneqq d(i, j^\prime )\), \(d_i^{\ast} = d(i, j^{\ast} )\), then for every \(S_{\ell}\), we have
	\begin{enumerate}[(a)]
		\item \(S_{\ell }\ni j^{\ast} \). Then contribution of \(i\) to \(\mathop{\mathrm{cost}}(Q^\prime ) - \mathop{\mathrm{cost}}(S_{\ell }\triangle Q^\prime)\) is at least \(d_i^\prime - d_i^{\ast}\).\footnote{\(i\) can go to \(j^{\ast} \).}
		\item \(S_{\ell }\ni j^\prime\). By the second property, either
		      \begin{itemize}
			      \item \(\pi (j^{\ast} )\in S_{\ell } \): this implies \(j^{\ast}\in S_{\ell }\), which falls back to the first case.
			      \item \(\pi (j^{\ast} )\notin S_{\ell } \): the contribution of \(i\) to \(\mathop{\mathrm{cost}}(Q^\prime ) - \mathop{\mathrm{cost}}(S_{\ell }\triangle Q^\prime)\) is at least \(d_i^\prime - (2d_i^{\ast} + d_i^\prime ) = -2d_i^{\ast}\).\footnote{\(i\) can go to \(\pi (j^{\ast} )\).}
		      \end{itemize}
		\item Otherwise, contribution of \(i\) to \(\mathop{\mathrm{cost}}(Q^\prime ) - \mathop{\mathrm{cost}}(S_{\ell }\triangle Q^\prime)\) is at least \(0\).\footnote{\(i\) can stay with \(j^\prime \).}
	\end{enumerate}

	In all, we see that the first case has total weight \(1\) from the first property, while (b)\(-\)(a) has total weight \(\leq p^\prime \), hence \autoref{eq:lec8} implies
	\[
		\sum_{i\in P} \left[ (d_i^\prime - d_i^{\ast} )\cdot 1 - (2d_i^{\ast} )\cdot p^\prime \right] \leq \sum_{\ell =1} ^t p_{\ell } (\mathop{\mathrm{cost}}(Q^\prime ) - \mathop{\mathrm{cost}}(S_{\ell }\triangle Q^\prime)) \leq 0,
	\]
	which is equivalent to say
	\[
		\mathop{\mathrm{cost}}(Q^\prime ) - (1 + 2p^\prime )\OPT \leq 0,
	\]
	so we get a \((1 + 2p^\prime )\)-approximation ratio. Furthermore, From \autoref{lma:lec8}, we have \(p^\prime = 1 + 2 / w\), hence we can achieve \(1 + 2(1 + 2 / w) = 3 + 4 / w\)-approximation ratio. Given \(\epsilon > 0\), by setting \(w \coloneqq 4/\epsilon \), we're done.
\end{proof}